Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-06-05 Grassmannian_cells.tex
Added Theorem 2.1, Lemma 2.2 and proof of 2.7

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\ui}{\underline i}
\newcommand{\uj}{\underline j}
\newcommand\udim{{\underline{\dim}\, }}

\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}

\title{Cell Decomposition of Rank 2 Quiver Grassmannians}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torus Action on Quiver Grassmannians}
\noindent Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:i\to j$. We denote by $\Rep(Q)$ the category of $\C$-representations of $Q$. Moreover, let $A_Q\cong \ZZ^{Q_1}$ be the free abelian group and $W_{Q}$ be the free non-abelian group generated by $Q_1$.  
\begin{definition}
The universal abelian covering quiver $\hat Q$ of $Q$ is given by the vertices $\hat Q_0=Q_0\times A_{Q}$ and the arrow set $\hat Q_1=Q_1\times A_{Q}$ where $(\alpha,\chi):(i,\chi)\to (j,\chi+e_\alpha)$ for every $\alpha:i\to j$.

@@

\end{proof}
In the following, we say that $d:\mathrm{supp}(X)\to\ZZ$ satisfies the degree condition for $X$ if it has the properties of Lemma \ref{degreecondition}.
\begin{theorem}Let $X$ be a representation which can be lifted to $\hat Q$ and choose $d:\mathrm{supp}(X)\to \ZZ$ such that it satisfies the degree condition for $X$. Then the torus action of $T$ on $\bigoplus_{q\in Q_0}X_q$ induced by $t.x_{(q,\chi)}=t^{d(q,\chi)}x_{(q,\chi)}$ for $x_{(q,\chi)}\in X_{(q,\chi)}$ induces
 a torus action on $Gr_\bfe^Q(X)$ such that
  $$Gr^Q_\bfe(X)^T\cong\bigsqcup_{[\hat\bfe]} Gr^{\hat Q}_{\hat\bfe}(\hat X)$$ where $\hat\bfe$ runs through all equivalence classes of dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
\end{proof}
The next step is to iterate the torus action keeping in mind the following idea: every representation, which lifts to the universal covering quiver, also lifts to the universal abelian covering quiver and also to the iterated universal abelian covering quivers, i.e. to $\hat Q^{(k)}:=\widehat{\hat Q^{(k-1)}}$ with $\hat Q=\hat Q^{(1)}$. Now it is straightforward to check that there exist natural surjective morphisms $f_k:\tilde Q\to \hat Q^{(k)}$ which become injective on finite subquivers if $k\gg 0$, see also \cite[Section 3.4]{wei}. Since the support of $X$ is finite as a representation of $\tilde Q$, we can find $k\geq 1$ such that the full subquiver with vertices $\mathrm{supp}(X)\subseteq \hat Q^{(k)}_0$ is a tree. Thus, if $\hat X^{(k)}$ is the lift of $X$ to $\hat Q^{(k)}$, there exists a torus action on the vector spaces $\hat X^{(k)}_q$ which induce a torus action on the quiver Grassmannian $Gr_{\bfe}^{\hat Q^{(k)}}(\hat X^{(k)})$. If we denote the iterated torus fixed points by $Gr^Q_\bfe(X)^{T^{(k)}}$. We obtain:
\begin{corollary}
Let $X$ be a representation which can be lifted to $\tilde Q$. Then there exists an iterated torus action such that
  $$Gr^Q_\bfe(X)^{T^{(k)}}\cong \bigsqcup_{[\hat\bfe^{(k-1)}]} Gr_{\hat\bfe^{(k-1)}}^{\hat Q^{(k-1)}}(\hat X^{(k-1)})^T\cong \bigsqcup_{[\hat\bfe^{(k)}]} Gr^{\hat Q}_{\hat\bfe^{(k)}}(\hat X^{(k)})\cong \bigsqcup_{[\tilde\bfe]} Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$$ where $\hat\bfe^{(k-1)},\,\hat\bfe^{(k)},\,\tilde\bfe$ run through all equivalence classes of dimension vectors compatible with $\bfe$.
\end{corollary}
\begin{corollary}
  affine bundles over $Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$ if $Gr^Q_\bfe(X)$ is smooth
  \[\{U\in Gr^Q_\bfe(X):\lim_{t\to0} t\cdot U\in Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)\}\]
\end{corollary}

\begin{question}
  What are the ranks of these bundles?  Poincar\'e polynomials?
\end{question}

@@

Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K(n)_0=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$. By $S_1$ and $S_2$ we denote the BGP-reflection functor corresponding to the sink and source of $K(n)$ respectively \cite{bgp}. Note that the quiver $K(n)$ itself does not change when applying the reflection functors, but the vertices $1$ and $2$ are exchanged. Let $$\mathcal S:=\begin{cases}S_1\text{ if }1 \text{ is a source }\\S_2\text{ if }$2$ \text{ is a source }\end{cases}$$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Pairs Label Cells in $Gr^Q_\bfe(P_m)$}
\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{cc}
Caldero, P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).

@@

\bibitem{sch} Schofield, A.: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).
\bibitem{wei} Weist, T.: Localization in quiver moduli spaces.
\end{thebibliography}

>>>>

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\ui}{\underline i}
\newcommand{\uj}{\underline j}
\newcommand\udim{{\underline{\dim}\, }}

\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\setcounter{MaxMatrixCols}{20}
\title{Cell Decomposition of Rank 2 Quiver Grassmannians}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torus Actions on Quiver Grassmannians}
\noindent The aim of this section is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general. If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bialynicki-Birula. More detailed, let $X$ be a smooth projective variety with a $\C^\ast$-action. For a connecting component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[\mathrm{Att}(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}\]

Then we have the following statement, \cite[Section 4]{bb}:
\begin{theorem}Let $\coprod_{i=1}^r C_i= X^{\C^\ast}$ be the decomposition into connected components. Then $\mathrm{Att}(C_i)$ is a locally closed smooth $\C^\ast$-invariant subvariety of $X$ where $C_i$ is a subvariety of $\mathrm{Att}(C_i)$. Moreover, we have $X=\coprod_{i=1}^r\mathrm{Att}(C_i)$ and
the natural map $\gamma_i:\mathrm{Att}(C_i)\to C_i$ is an affine vector bundle.

In particular, we have $\chi(X)=\chi(X^{\C^\ast})$ for the Euler characteristic.
\end{theorem}

Fix a vector space $X$ of dimension $n$ and let $d\leq n$. We first consider a natural torus action on usual Grassmannians coming along with any direct sum decomposition of the vector space $X$. Then we generalize the concept to quiver Grassmannians, but we will see that the torus fixed point set can be calculated in an analogous manner. Thus assume that $X=\bigoplus_{j=1}^l X_j$ is a direct sum decomposition of $X$. We choose a basis $\mathcal B=\{v_1,\ldots,v_n\}$ of $X$ which is compatible with this decomposition, i.e. there exist number $1\leq j_1\leq j_2\ldots\leq j_{l-1}\leq n$ such that
\[v_1,\ldots,v_{j_1}\in X_1,\,v_{j_1+1},\ldots,v_{j_2}\in X_2,\ldots,v_{j_{l-1}+1},\ldots,v_{n}\in X_l\]
Moreover, we choose a map $d:\{1,\ldots,n\}\to\mathbb{Z}$ such that $d(i)=d(i')$ if $v_i,v_{i'}\in X_{j}$ for some $j$. Then there exists a $\C^\ast$-action on $X$ when extending the definition $t.v_i:=t^{d(j)}v_i$ if $v_i\in X_j$ linearly to $X$. This naturally induces an action of $\C^\ast$ on the usual Grassmannians $\Gr_d(X)$. We can write every $U\in\Gr_d(X)$ uniquely in row-echelon form with respect to the basis $\mathcal B$, i.e. there exists a unique sequence $1\leq i_1\leq\ldots\leq i_d\leq n$ and a unique matrix representation of $U$ of the form
\[M(U):=\begin{pmatrix}\ast&\ldots &\ast &1&0 &\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&0\\\ast&\ldots &\ast&0&\ast&\ldots &\ast&1&0&\ldots&\ldots&\ldots&\ldots&\ldots&0\\\vdots &\ddots&\vdots&\vdots&\vdots &\ddots&\vdots&0&\ast&\ddots&\ddots&\ddots&\ddots&\ddots&\vdots\\\vdots &\ddots&\vdots&\vdots&\vdots &\ddots&\vdots&\vdots&\ddots&\ddots&\ddots&\ddots&\ddots&\ddots&\vdots\\\ast&\ldots &\ast&0&\ast&\ldots &\ast&0&\ast&\ldots&\ast&1&0&\ldots&0\end{pmatrix}\in M_{d,n}(k)\]
where the unit vectors are in the columns $i_1,\ldots,i_d$. All matrix representations of this form define a Schubert cell $X_{\bfi}$ where $\bfi=(i_1,\ldots,i_d)$. The $\C^\ast$-action is then given by multiplying the $w^{\mathrm{th}}$ column with $t^{d(w)}$ where $v_w\in X_j$, i.e. we have
\[(t.M(U))_{u,w}=t^{d(w)}\cdot M(U)_{u,w}.\]
Every Schubert cell is invariant under the torus action, i.e. $t.M(U)\in X_\bfi$ for all $M(U)\in X_{\bfi}$. 
\begin{lemma}\label{usualGrass}
Choose $d$ in such a way that $d(i)\neq d(i')$ if $v_i\in X_j$, $v_{i'}\in X_{j'}$ and $j\neq j'$.
Then we have $U\in X_{\bfi}^{\C^\ast}$ if and only if $U=\bigoplus_{i=1}^l U\cap X_i$.
\end{lemma}
\begin{proof}If $U=\bigoplus_{i=1}^l U\cap X_i$, we have 
\[t.U=\bigoplus_{i=1}^l t.(U\cap X_i)=\bigoplus_{i=1}^l t^{d(j_i+1)}(U\cap X_i)=U\]
where $j_0=0$.

Assume that $U$ is a torus fixed with $M(U)\in X_{\bfi}$. Fix the $w^{\mathrm{th}}$ column $M(U)_w$ and assume that $v_{i_w}\in X_s$. As the matrix representation $M(U)$ is uniquely determined, by the assumption on $d$, it follows that all matrix coefficients corresponding to a basis element $v_i$ such that $v_i\notin X_s$ are zero. This means that $\widehat{M(U)}_w\in X_s$ where $\widehat{M(U)}_w$ is the vector of $U$ corresponding to $M(U)_w$. As the colums are linear independent, this already shows the claim.
\end{proof}
The next step is to generalize this to quiver Grassmannians.
Now let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:i\to j$. We denote by $\Rep(Q)$ the category of $\C$-representations of $Q$. Moreover, let $A_Q\cong \ZZ^{Q_1}$ be the free abelian group and $W_{Q}$ be the free non-abelian group generated by $Q_1$.  
\begin{definition}
The universal abelian covering quiver $\hat Q$ of $Q$ is given by the vertices $\hat Q_0=Q_0\times A_{Q}$ and the arrow set $\hat Q_1=Q_1\times A_{Q}$ where $(\alpha,\chi):(i,\chi)\to (j,\chi+e_\alpha)$ for every $\alpha:i\to j$.

@@

\end{proof}
In the following, we say that $d:\mathrm{supp}(X)\to\ZZ$ satisfies the degree condition for $X$ if it has the properties of Lemma \ref{degreecondition}.

\begin{theorem}Let $X$ be a representation which can be lifted to $\hat Q$ and choose $d:\mathrm{supp}(X)\to \ZZ$ such that it satisfies the degree condition for $X$. Then the torus action of $T$ on $\bigoplus_{q\in Q_0}X_q$ induced by $t.x_{(q,\chi)}=t^{d(q,\chi)}x_{(q,\chi)}$ for $x_{(q,\chi)}\in X_{(q,\chi)}$ induces
 a torus action on $Gr_\bfe^Q(X)$ such that
  $$Gr^Q_\bfe(X)^T\cong\bigsqcup_{[\hat\bfe]} Gr^{\hat Q}_{\hat\bfe}(\hat X)$$ where $\hat\bfe$ runs through all equivalence classes of dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
A representation $U\in\Gr_{\bfe}(X)$ is a torus fixed point if and only if $t.U=U $ for all $t\in\C^\ast$, i.e. $t.U_q=U_q$ for all $q\in Q_0$, $t\in\C^\ast$. Thus, apart from being a subrepresentation of $X$, it is a fixed point of the torus actions on the usual Grassmannians $\Gr_{\bfe_q}(X_q)$ induced by the decompositions $X_q=\bigoplus_{\chi\in A_Q} X_{q,\chi}$. A subrepresentation $U\subset X$ can be lifted to the universal abelian covering if and only if
\[X_q\cap U_q=\bigoplus_{\chi\in A_Q} X_{(q,\chi)}\cap U_{(q,\chi)}=\bigoplus_{\chi\in A_Q}U_{(q,\chi)}.\]
Thus we can apply Lemma \ref{usualGrass} to obtain the claim.
\end{proof}
The next step is to iterate the torus action keeping in mind the following idea: every representation, which lifts to the universal covering quiver, also lifts to the universal abelian covering quiver and also to the iterated universal abelian covering quivers, i.e. to $\hat Q^{(k)}:=\widehat{\hat Q^{(k-1)}}$ with $\hat Q=\hat Q^{(1)}$. Now it is straightforward to check that there exist natural surjective morphisms $f_k:\tilde Q\to \hat Q^{(k)}$ which become injective on finite subquivers if $k\gg 0$, see also \cite[Section 3.4]{wei}. Since the support of $X$ is finite as a representation of $\tilde Q$, we can find $k\geq 1$ such that the full subquiver with vertices $\mathrm{supp}(X)\subseteq \hat Q^{(k)}_0$ is a tree. Thus, if $\hat X^{(k)}$ is the lift of $X$ to $\hat Q^{(k)}$, there exists a torus action on the vector spaces $\hat X^{(k)}_q$ which induce a torus action on the quiver Grassmannian $Gr_{\bfe}^{\hat Q^{(k)}}(\hat X^{(k)})$. If we denote the iterated torus fixed points by $Gr^Q_\bfe(X)^{T^{(k)}}$. We obtain:
\begin{corollary}
Let $X$ be a representation which can be lifted to $\tilde Q$. Then there exists an iterated torus action such that
  $$Gr^Q_\bfe(X)^{T^{(k)}}\cong \bigsqcup_{[\hat\bfe^{(k-1)}]} Gr_{\hat\bfe^{(k-1)}}^{\hat Q^{(k-1)}}(\hat X^{(k-1)})^T\cong \bigsqcup_{[\hat\bfe^{(k)}]} Gr^{\hat Q}_{\hat\bfe^{(k)}}(\hat X^{(k)})\cong \bigsqcup_{[\tilde\bfe]} Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$$ where $\hat\bfe^{(k-1)},\,\hat\bfe^{(k)},\,\tilde\bfe$ run through all equivalence classes of dimension vectors compatible with $\bfe$.
\end{corollary}
%\begin{corollary}
%  affine bundles over $Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$ if $Gr^Q_\bfe(X)$ is smooth
%  \[\{U\in Gr^Q_\bfe(X):\lim_{t\to0} t\cdot U\in Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)\}\]
%\end{corollary}

\begin{question}
  What are the ranks of the bundles over the fixed point set?  Poincar\'e polynomials?
\end{question}

@@

\noindent Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K(n)_0=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$. By $S_1$ and $S_2$ we denote the BGP-reflection functor corresponding to the sink and source of $K(n)$ respectively \cite{bgp}. Note that the quiver $K(n)$ itself does not change when applying the reflection functors, but the vertices $1$ and $2$ are exchanged. Let $$\mathcal S:=\begin{cases}S_1\text{ if }1 \text{ is a source }\\S_2\text{ if }$2$ \text{ is a source }\end{cases}$$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Pairs Label Cells in $Gr^Q_\bfe(P_m)$}
\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{cc}
Caldero, P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).

@@

\bibitem{sch} Schofield, A.: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).
	\bibitem{wei} Weist, T.: Localization of quiver moduli spaces. Representation Theory \textbf{17}(13), 382-425 (2013).
\end{thebibliography}
